The Resurgence of the Linear Optics Interferometer --- Recent Advances \& Applications

### 2017-04-27 paper.tex
Updated single photon.

In SPDC, a nonlinear crystal with a large $\chi^{(2)}$ is pumped with a strong 
laser and with a small probability, the pump beam is absorbed by the crystal to 
produce two beams of lower energy known as the signal and idler beams. Owing to 
conservation of energy and momentum, the two beams have spatio-temporal 
correlations that can be engineered to produce twin-beam states which have 
perfectly correlated photon numbers. If a single photon were to be detected in one 
of the beams, and then it is certain that the other beam would be in the state of a 
single photon. Some commonly used nonlinear crystals for SPDC are beta barium 
borate (BBO), periodically poled lithium niobate (PPLN), and periodically poled 
potassium titanyl phosphate (PPKTP). Recently, techniques using bismuth triborate 
(BiBO) have improved to the extent of becoming one of the record-holders of single-
photon production.
\cite{bib:WangChen16}.

>>>>

In SPDC, a nonlinear crystal with a large $\chi^{(2)}$ is pumped with a strong 
laser and with a small probability, the pump beam is absorbed by the crystal to 
produce two beams of lower energy known as the signal and idler beams. Owing to 
conservation of energy and momentum, the two beams have spatio-temporal 
correlations that can be engineered to produce twin-beam states which have 
perfectly correlated photon numbers. If a single photon were to be detected in one 
of the beams, and then it is certain that the other beam would be in the state of a 
single photon. Some commonly used nonlinear crystals for SPDC are beta barium 
borate (BBO), periodically poled lithium niobate (PPLN), and periodically poled 
potassium titanyl phosphate (PPKTP). Recently, techniques using bismuth triborate 
(BiBO) have improved to the extent of becoming one of the record-holders of single-
photon production \cite{bib:WangChen16}.

\sihui{ Discuss architectures for producing entangled photon states like GHZ, maybe here or in a different subsection.}

Solid-state single photon sources come from semiconducting nanostructures and 
nitrogen valencies (NV) centers in diamond. Both types are versatile and efficient 
sources of single photons, however, the photons they produce have suffered from 
lack of indistinguishability that is necessary for quantum information processing.
Recent developments in the former show promising developments. Notably, much 
progress has been made using resonant excitation of quantum dots. Laser pulses are 
used to excite their electronic resonance and trigger the emission of single 
photons \cite{bib:Muller07,bib:Vamivakas09, bib:Flagg09,bib:Ates09, bib:Dirk10,
bib:1748,bib:Jayakumar13,bib:Wei14,bib:Muller14,bib:Unsleber15,
bib:237403,bib:Sweeney14}. An architecture that embeds the quantum dot in a 
micropillar cavity with the same resonant frequency as the dot exploits the 
Purcell effect to achieve higher single-photon production efficiency 
†\cite{bib:Ding16,bib:213601}. 
Such resonant excitation of quantum dots overcome the homogenous broadening of the 
excited state that causes degradation of photon purity and hence improves
indistinguishability. Using a time-bin loop-based architecture \cite{bib:Motes14} 
on a stream of single photons produced by such a quantum dot, the same group was 
able to demonstrate quantum interference between five indistinguishable photons 
\cite{bib:WangChen16}.
